---
title: "Programming Assignment"
subtitle: "4"
author: "Yhosep Barba"
institute: "Rutgers University"
date: "2025/03/31 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
#Uploading packages
```{r}
library("tidyverse")
library("here")
library("ds4ling")
```

```{r}
library("knitr")
```

---
#Uploading datasets
```{r}
data1 <- read_csv(here("data_raw", "fricatives", "assumptions.csv"))
data2 <- read_csv(here("data_raw", "fricatives", "fricatives.csv"))  
```

---
#Tidying data
```{r}
#| echo = FALSE
data2_tidy <- data2 |>
  pivot_longer(
    cols = c(s_cog, s_skewness, sh_cog, sh_skewness),
    names_to = "measurement",
    values_to = "value"
  ) |>
  separate(
    col = measurement,
    into = c("phoneme", "measurement"),
    sep = "_"
  ) |>
   pivot_wider(
    names_from = measurement,
    values_from = value                       
  ) 
print(data2_tidy)
```

---
#Table of descriptive statistics
```{r}
#| echo = FALSE
summary_table <- data2_tidy |>
  group_by(phoneme) |>
  summarise(
    mean_skew = mean(skewness),
    mean_cog = mean(cog),
    min_skew = min(skewness),
    min_cog = min(cog),
    mead_skew = median (skewness),
    mead_cog = median (cog)
  )

knitr::kable(summary_table)
```

---
#Boxplot of Center of Gravity by Phoneme
```{r}
#| echo = FALSE
data2_tidy |>
  ggplot() +
  aes(x = phoneme, y = cog) +
  geom_boxplot() +
  labs(
    title = "Boxplot of Center of Gravity by Phoneme",
    x = "Phoneme",
    y = "Center of Gravity"
  )
```

---
#Skewness as a function of phoneme
```{r}
#| echo = FALSE
data2_tidy |>
  ggplot() +
  aes(x = phoneme, y = skewness) +
  stat_summary(
    fun.data = "mean_se",  
    geom = "point",        
    size = 3,              
    color = "red"         
  ) +
  labs(
    title = "Skewness by Phoneme",
    x = "Phoneme",
    y = "Skewness"
  )
```
---
#Center of gravity as a function of skewness for the [s]
```{r}
s_data <- data2_tidy |>
  filter(phoneme == "s") 

model <- lm(cog ~ skewness, data = s_data)

print(model)
```
---
#Scatter Plot
```{r}
#| echo = FALSE
s_data |>
  ggplot() +
  aes(x = skewness, y = cog) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Center of Gravity as a function of Skewness for [s] Segments",
    x = "Skewness",
    y = "Center of Gravity"
  )

```
---
#Model Table
```{r}
knitr::kable(summary(model)$coef)
```

---
#Model Diagnosis
```{r}
diagnosis(model)
```

---
#Results
The data were analyzed using a linear regression model to examine the relationship between skewness and center of gravity (cog) for the [s] phoneme. In this model, I used skewness as a predictor variable and the center of gravity as the dependent variable.The analysis showed a significant positive relationship between skewness and the center of gravity. Specifically, the coefficient for skewness was 1675.5933, indicating that for each unit increase in skewness, the center of gravity increased by approximately 1675.6 units. The intercept was -756.6888, suggesting that when skewness is zero, the center of gravity is approximately -756.7 units. The p-value for skewness was less than 0.0001, which indicates that this relationship is significant.

The results were consistent across diagnostic checks, suggesting a  relationship between skewness and center of gravity for the [s] phoneme.

---

```{r}
#| echo = FALSE
data1|>
  ggplot() +
  geom_point() +
  aes (x = x, y = y) +
  labs (x = "x", y = "y", title = "Assumptions")
```
##There is no real data. There is a dinosaur
---








